\myTitle

### 2023-10-25 body-text-of-paper.tex
A number of edits
Added permutations figure

	In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of mixed ancestry--- 
	characterized as individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference.}
	a trait often summarized as the proportion of an individual’s genome originating from each of $K$ subpopulations, sensu \citet{pritchard2000inference}.  Various processes could contribute to such mixed associations.  Most directly, individuals may represent the influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately representing contributions from multiple sampled populations.  However, similar patterns may result from functionally different patterns of population structuring.  For example, such patterns might be expected when clinal systems governed by isolation by distance are described as discrete genetic clusters. 
	NOTE I think a line here about exogenous immigrants might be worthwhile to set up 2 biological processes paired with a statistical process, something like END NOTE
	Alternatively, individuals of mixed ancestry could reflect immigration from outlying populations (e.g., exogenous immigration), with true source populations not sufficiently
	represented in the sample to be among the characterized $K$ subpopulations.
	Furthermore, the characterization of admixed individuals may reflect a limitation of the statistical power of the genetic data, 
	as a function of both the marker set and sample size, to resolve the true underlying patterns of genetic structure. 
	 The challenge for researchers then becomes correctly identifying the ecological and/or statistical processes by which observed complex ancestry patterns were created.

@@

	In response to these needs, we developed gscramble – a simulation approach that samples alleles from respective populations without replacement, thus maintaining the genetic differences in the empirical dataset.
	Further, by simulating individual genotypes based on user-defined pedigrees, gscramble allows for the simulation of admixed genotypes of varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
	By integrating species-specific recombination rates, gscramble simulates biologically informed genotypes that mirror empirical data.
	We develop and illustrate gscramble with use of single nucleotide polymorphic (SNP) genotypes.

@@

Before we proceed to a description of the methodology used by gscramble,
we demonstrate a case in which using a standard sampling-with-replacement
approach induces RIPI when estimating admixture fractions.
To construct this simple example, we used the R programming language \citep{Rcoreteam} to
simulate allele frequencies for $L$ biallelic
loci in a single population from a $\mathrm{Beta}(1, 8)$ distribution and then simulated
two samples, $A$ and $B$, each of size $N$ diploid genotypes. The genotypes in $A$ and $B$ were simulated
from the same allele frequencies assuming Hardy-Weinberg equilibrium.

In this case,
each set of $N$ individuals is a sample from exactly the same population, so
there is no basis for performing population assignment between $A$ and $B$;
however, we treated samples $A$ and $B$ as if they were sampled from two potentially different
populations, and simulated $9n$ new individuals: $n$ new individuals at each of the 9 values of
$q_A$, the admixture fraction for population $A$, in  $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$. We simulated these genotypes according to the "admixture model" used in {\em structure} \citep{pritchard2000inference}.  Briefly, at each locus, independently, the origin
of each gene copy was simulated to be from
population $A$ with probability $q_A$ and from population $B$ with probability $1-q_A$.  Subsequently,
the allelic types of gene copies from $A$ ($B$), at a locus, were simulated by sampling with replacement
from the alleles at that locus in sample $A$ ($B$).
We then added these $9n$ simulated individuals to the $N$ original individuals  from $A$ and $N$ from $B$
and analyzed all the genotypes using ADMIXTURE with $K=2$. On each data set we performed a supervised analysis in which the original $N$ samples from each
population were provided as learning samples, and also an
unsupervised analysis when the origin of the samples from $A$ and $B$ were regarded as unknown.

@@

Our approach for simulating admixed individuals without inflating perceived power for
the estimation of their admixture fractions is motivated by the traditional cross-validation
approach for evaluating the power of mixture proportion
estimation in genetic stock identification \citep{Moran&AndersonRubias}.
In that cross-validation approach, individuals are removed from the reference
samples and placed in a mixture sample of "unknowns", whose origin
is inferred using the remaining individuals in the reference samples.  A key
feature of this approach is that no new genetic material is being created by
sampling with replacement; however the genetic material of all individuals
in the original samples is being used---either within the remaining individuals in the
reference samples or within the individuals in the mixture. This latter condition
is not essential, but it does indicate that the methodology is using all the
available individuals for estimating power through simulation.

Our scheme extends the traditional cross-validation approach
by simulating the descent of chunks of chromosome through a user-specified
pedigree to create admixed individuals.  By imposing certain constraints
on the pedigree, our framework ensures that all the genetic material
within the original individuals is used to either create admixed individuals or to
remain within reference individuals (i.e., those that are purely of one
putative subopulation), and yet no new genetic material is
created by sampling with replacement from the original samples.  We call the structure
we use to do this a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
a GSP includes a set of founders (individuals that have no parents specified in the
pedigree), and it includes non-founders; however, we introduce an additional node
type in the GSP that represents samples taken from the non-founders. These samples
are the simulated admixed individuals.  A GSP must have no inbreeding loops, since
any inbreeding loop indicates that more than one copy of a gene in an ancestor has
been created amongst its descendants, which is a type of sampling without
replacement.  Furthermore, in a GSP, the amount of genetic material taken from the
samples at the sample nodes should be equal to the amount of material in the founders,
to ensure that all available data is being used to estimate power for the estimation of
admixture fractions.

Figure~\ref{fig:gsp1} shows an example GSP for the simple case
of simulating $F_2$ individuals. The figure and its caption
should be studied, as the following section
uses terminology that is defined in the figure caption.
%%%%%%%%%%%
\begin{figure}
\newcommand{\gspcapone}{\footnotesize 
(I NEED TO UPDATE THE FIGURE TO MATCH THE DESCRIPTION AND MAKE
THE LABELS BIGGER IN THE SQUARES).
A genomic simulation pedigree (GSP) for simulating $F_2$ individuals.
Squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
individuals 1--4.  The two triangles above each of those founders denotes, in color
and text, the population of origin of each of the two genomes within each founder.
Thus, individuals~1 and~2 are founders purely from population~$A$, and~3 and~4
are purely from population~$B$.  The pink hexagon labeled s7 is the sample node.
It represents samples that are taken from non-founder node~7, which, from the structure
of the pedigree, clearly represents an $F_2$ hybrid between populations $A$ and $B$.  The red numbers
on edges above and leading into each non-founder node indicate the number of
gametes that get segregated from the parent node (at the top of the edge) to the
daughter node (at the bottom of the edge).
This emphasizes that,
in dropping genes through a GSP, in contrast
to a traditional pedigree, each parent
node may segregate multiple gametes to a daughter node.  See the text for
further explanation. 
Each non-founder node $x$ will have exactly two edges coming into it from above
(from its two parents), we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$. 
For example, node 6 has $G^+_{6,1}=G^+_{6,2} = 2$.
A non-founder individual may have more or less than two edges coming downward
out of it.  Letting $e_x$ be the number of downward edges out of node $x$ to its $e_x$
non-founder daughter nodes, we denote the number of gametes segregated
through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example,
non-founder node~5, $e_5 = 1$ and $G^{-}_{5,1}=4$.
Finally, the purple number adjacent to the edge leading into the
sample node indicates how many sample individuals are drawn from individual
node 7. Any non-founder individual node can have a maximum of one edge leading
to a sample node.  We denote the number of samples emanating from node $x$
node as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
created from a single simulation on this GSP.
}
\begin{center}
\includegraphics[width=\columnwidth]{images/gsp4-700.pdf}
\end{center}
\caption[\gspcapone]{\gspcapone}
\label{fig:gsp1}
\end{figure}
%%%%%%%%%%%%
It is important to understand that, apart from the founder nodes, the individual
nodes in a GSP do not necessarily represent only a single individual.  Specifically,
since we are preserving all genetic material, multiple gametes may get segregated
through each non-founder node in a GSP\@.   Genetic material is segregated through
a GSP using these steps, done upon each node in an order such that the steps
are run on each node's parents before being run on the node itself:
\begin{enumerate}
\item Each genome in a triangle above a founder node represents a gamete.
These gametes are united within the founder individual and then
following recombination between each chromosome
with probability $\frac{1}{2}$ and within each chromosome according to
a user-specified recombination map, 
 one gamete is assorted down each edge from the founder node to its daughter.
\item Each non-founder node, $x$, will have exactly two edges coming
into it---one from each parent.  Each of the $G^+_{x,1}$ gametes coming in from the
first edge is united with a single one of the $G^+_{x,2}$ gametes coming
in on the second edge (in a random order). 
\item If a non-founder node $x$ has an edge to a sample node,
a randomly chosen $S_x$ of the united pairs of gametes are delivered
to the sample node to constitute the $S_x$ samples taken from individual
node $x$. Any remaining united pairs of gametes are treated as in (4) below.
\item United pairs of gametes (including those remaining after delivery
to a sample node) in each non-founder node $x$ are segregated into
gametes with recombination, and each of these gametes is assigned, in
random order, to the gamete edges proceeding downward out of the node,
with the number of gametes assigned to the $i\thh$ downward edge being
$G^-_{x,i}$. 
\end{enumerate}
At the end of this process, the samples simulated (as united pairs of gametes)
into the sample nodes will contain all of the genetic material found in the founders (and
no more than that)
but that material will have been segregated and recombined according to the pedigree.  Thus the
samples obtained from a GSP can be simulated with the expected admixture fractions
for any hybrid category that can be specified by a pedigree.  They have also been simulated
in a way that 1) respects physical linkage---each individual inherits segments of the genome
simulated using a recombination map; 2) does not duplicate any genetic material---the sampling of
genetic material is entirely without replacement so it will not induce RIPI; and 3) no genetic material has been lost---all the genetic material of the founders
is represented in the samples, maximizing the number of individuals and markers available to
use in downstream analyses.

In the gscramble package there is a convenience function called {\footnotesize\tt create\_GSP()} that
will create GSPs for samples up to and including second-generation backcrosses; however, gscramble is designed
to handle user-specified GSPs of arbitrary complexity.  When creating one's own GSP, it is useful
to keep in mind a handful of requirements for a GSP to be valid. Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder nodes, and $\nonfounds$ for all non-founder
nodes that are not parents of other non-founders, but are adjacent to a sample node, $\nonfounde$ for all
non-founder nodes that have non-founder node daughters, but no sample nodes, and $\nonfoundse$ for all
non-founder nodes that are parents of non-founder nodes and also adjacent to sample-nodes, the necessary conditions 
for a valid GSP, in addition to having no inbreeding loops, are:
\begin{equation}
\begin{aligned}
\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders| &  \\ 
\sum_{i=1}^{e_x} G^-_{x,i} &= 2 & \forall x \in \founders \\
G^+_{x,1} &= G^+_{x,2} & \forall x \in \nonfound \\
\sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\ 
 S_x/2 + \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfoundse.  \\ 
\end{aligned}
\end{equation}
Before simulating from a GSP, gscramble checks to ensure the GSP
satisfies these validity conditions.

@@

gscramble uses a GSP to segregate segments of genetic material from the founders
to the samples.  At the end of segment simulation, the genome of each sample
is recorded as a mosaic of segments that derived from different founders.  In order
to create a data set of markers in each sample, it is necessary only to identify the alleles
that occurred on each segment within the founders and propagate those alleles to the
corresponding segments within the samples.  At the stage of assigning alleles to
founders, gscramble provides a variety of options for to permute alleles amongst
individuals within samples from which the founders are drawn.  By using permutation,
variability in the data is created, but without sampling with replacement, and hence
not inducing RIPI.

@@

GSCRAMBLE REFERENCE GROUPS
GSCRAMBLE PEDIGREES
ADMIXTURE GSCRAMBLE VS EMPIRICAL
FST OF EMPIRICAL AND GSCRAMBLE CORE POPULATIONS

>>>>

	In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of mixed ancestry---
	characterized as individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference}.
	a trait often summarized as the proportion of an individual’s genome originating from each of $K$ subpopulations, sensu \citet{pritchard2000inference}.  Various processes could contribute to such mixed associations.  Most directly, individuals may represent the influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately representing contributions from multiple sampled populations.  However, similar patterns may result from functionally different patterns of population structuring.  For example, such patterns might be expected when clinal systems governed by isolation by distance are described as discrete genetic clusters.
	NOTE I think a line here about exogenous immigrants might be worthwhile to set up 2 biological processes paired with a statistical process, something like END NOTE
	Alternatively, individuals of mixed ancestry could reflect immigration from outlying populations (e.g., exogenous immigration), with true source populations not sufficiently
	represented in the sample to be among the characterized $K$ subpopulations.
	Furthermore, the characterization of admixed individuals may reflect a limitation of the statistical power of the genetic data,
	as a function of both the marker set and sample size, to resolve the true underlying patterns of genetic structure.
	 The challenge for researchers then becomes correctly identifying the ecological and/or statistical processes by which observed complex ancestry patterns were created.

@@

	In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without replacement, thus maintaining the genetic differences in the empirical dataset.
	Further, by simulating individual genotypes based on user-defined pedigrees, \gscramble{} allows for the simulation of admixed genotypes of varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
	By integrating species-specific recombination rates, \gscramble{} simulates biologically informed genotypes that mirror empirical data.
	We develop and illustrate \gscramble{} with use of single nucleotide polymorphic (SNP) genotypes.

@@

Before we proceed to a description of the methodology used by \gscramble{},
we demonstrate a case in which using a standard sampling-with-replacement
approach induces RIPI when estimating admixture fractions.
To construct this simple example, we used the R programming language \citep{Rcoreteam} to
simulate allele frequencies for $L$ biallelic
loci in a single population from a $\mathrm{Beta}(1, 8)$ distribution and then simulated
two samples, $A$ and $B$, each of size $N$ diploid genotypes. The genotypes in $A$ and $B$ were simulated
from the same allele frequencies assuming Hardy-Weinberg equilibrium.

In this case,
each set of $N$ individuals is a sample from exactly the same population, so
there is clearly no basis for performing population assignment between $A$ and $B$;
however, we treated samples $A$ and $B$ as if they were sampled from two potentially different
populations, and simulated $9n$ new individuals: $n$ new individuals at each of the 9 values of
$q_A$, the admixture fraction for population $A$, in  $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$. We simulated these genotypes according to the "admixture model" used in {\em structure} \citep{pritchard2000inference}.  Briefly, at each locus, independently, the origin
of each gene copy was simulated to be from
population $A$ with probability $q_A$ and from population $B$ with probability $1-q_A$.  Subsequently,
the allelic types of gene copies from $A$ ($B$), at a locus, were simulated by sampling with replacement
from the alleles at that locus in sample $A$ ($B$).
We then added these $9n$ simulated individuals to the $N$ original individuals  from $A$ and $N$ from $B$
and analyzed all the genotypes using ADMIXTURE with $K=2$. On each data set we performed a supervised analysis in which the original $N$ samples from each
population were provided as learning samples, and also an
unsupervised analysis when the origin of the samples from $A$ and $B$ were regarded as unknown.

@@

Our approach for simulating admixed individuals without inflating perceived power for
the estimation of their admixture fractions is motivated by the traditional cross-validation
approach for evaluating the power of mixture proportion
estimation in genetic stock identification (e.g., the cross-validation approach
desribed in \citealt{moran2019bayesian}).
In the cross-validation approach, individuals are removed from the reference
samples and placed in a mixture sample of "unknowns", whose origin
is inferred using the remaining individuals in the reference samples.  A key
feature of this approach is that no new genetic material is being created by
sampling with replacement; however the genetic material of all individuals
in the original samples is being used---either within the remaining individuals in the
reference samples or within the individuals in the mixture.

Our scheme, described here, extends the traditional cross-validation approach
by simulating the descent of chunks of chromosome through a user-specified
pedigree to create admixed individuals.  By imposing certain constraints
on the pedigree, our framework ensures that all the genetic material
within the original individuals is used to either create admixed individuals or to
remain within reference individuals (i.e., those that are purely of one
putative subopulation), and yet no new genetic material is
created by sampling with replacement from the original samples.

The call the structure
we use to do this sort of cross-validation simulation of admixed individuals
a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
a GSP includes a set of founders (individuals that have no parents specified in the
pedigree), and it includes non-founders; however, we introduce an additional node
type in the GSP that represents samples taken from the non-founders. These samples
are the simulated admixed individuals.  A GSP must have no inbreeding loops, since
any inbreeding loop indicates that more than one copy of a gene in an ancestor may
be created amongst its descendants, which is a type of sampling without
replacement.  Furthermore, in a GSP, the amount of genetic material taken from the
samples at the sample nodes should be equal to the amount of material in the founders,
to ensure that all available data is being used to estimate power for the estimation of
admixture fractions.

Figure~\ref{fig:gsp1} shows an example GSP for the simple case
of simulating $F_2$ individuals. The figure and its caption
should be studied, as the following section
uses terminology that is defined in the figure caption.
%%%%%%%%%%%
\begin{figure}
\newcommand{\gspcapone}{\footnotesize
A genomic simulation pedigree (GSP) for simulating $F_2$ individuals.
Squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
individuals 1--4.  The inverted triangles above each of those founders denotes, in color
and text, the population of origin of each of the two genomes within each founder.
Thus, individuals~1 and~2 are founders purely from population~$A$, and~3 and~4
are purely from population~$B$.  The pink hexagon labeled s7 is the sample node.
It represents samples that are taken from non-founder node~7, which, from the structure
of the pedigree, clearly represents an $F_2$ hybrid between populations $A$ and $B$.  The red numbers
on edges above and leading into each non-founder node indicate the number of
gametes that get segregated from the parent node (at the top of the edge) to the
daughter node (at the bottom of the edge).
This emphasizes that,
in dropping genes through a GSP, in contrast
to a traditional pedigree, each parent
node may segregate multiple gametes to a daughter node.  See the text for
further explanation.
Each non-founder node $x$ will have exactly two edges coming into it from above
(from its two parents), we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$.
For example, node 6 has $G^+_{6,1}=G^+_{6,2} = 2$.
A non-founder individual may have more or less than two edges coming downward
out of it.  Letting $e_x$ be the number of downward edges out of node $x$ to its $e_x$
non-founder daughter nodes, we denote the number of gametes segregated
through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example,
non-founder node~5, $e_5 = 1$ and $G^{-}_{5,1}=4$.
Finally, the purple number adjacent to the edge leading into the
sample node indicates how many sample individuals are drawn from individual
node 7. Any non-founder individual node can have a maximum of one edge leading
to a sample node.  We denote the number of samples emanating from node $x$
node as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
created from a single simulation on this GSP.
}
\begin{center}
\includegraphics[width=\columnwidth]{figures/F2.pdf}
\end{center}
\caption[\gspcapone]{\gspcapone}
\label{fig:gsp1}
\end{figure}
%%%%%%%%%%%%
It is important to understand that, apart from the founder nodes, the individual
nodes in a GSP do not necessarily represent only a single individual.  Specifically,
since we are preserving all genetic material, multiple gametes may get segregated
through each non-founder node in a GSP\@.

Genetic material is segregated through
a GSP using these steps, done upon each node in an order such that the steps
are run on each node's parents before being run on the node itself:
\begin{enumerate}
\item Each genome in a triangle above a founder node represents a gamete
carrying a single, haploid copy of the genome.
These gametes are united within the founder individual and then
following recombination between each chromosome
with probability $\frac{1}{2}$ and within each chromosome according to
a user-specified recombination map,
 one gamete is assorted down each edge from the founder node to its daughter.
\item Each non-founder node, $x$, will have exactly two edges coming
into it---one from each parent.  Each of the $G^+_{x,1}$ gametes coming in from the
first edge is united with a single one of the $G^+_{x,2}$ gametes coming
in on the second edge (in a random order).
\item If a non-founder node $x$ has an edge to a sample node,
a randomly chosen $S_x$ of the united pairs of gametes are delivered
to the sample node to constitute the $S_x$ samples taken from individual
node $x$. Any remaining united pairs of gametes are treated as in (4) below.
\item United pairs of gametes (including those remaining after delivery
to a sample node) in each non-founder node $x$ are segregated into
gametes with recombination, and each of these gametes is assigned, in
random order, to the gamete edges proceeding downward out of the node,
with the number of gametes assigned to the $i\thh$ downward edge being
$G^-_{x,i}$.
\end{enumerate}
At the end of this process, the samples simulated (as united pairs of gametes)
into the sample nodes will contain all of the genetic material found in the founders (and
no more than that)
but that material will have been segregated and recombined according to the pedigree.  Thus the
samples obtained from a GSP can be simulated with the expected admixture fractions
for any hybrid category that can be specified by a pedigree.  They have also been simulated
in a way that 1) respects physical linkage---each individual inherits segments of the genome
simulated using a recombination map; 2) does not duplicate any genetic material---the sampling of
genetic material is entirely without replacement so it will not induce RIPI; and 3) no genetic material has been lost---all the genetic material of the founders
is represented in the samples, maximizing the number of individuals and markers available to
use in downstream analyses.

In the \gscramble{} package there is a convenience function called {\footnotesize\tt create\_GSP()} that
will create GSPs for samples up to and including second-generation backcrosses; however, \gscramble{} is designed
to handle user-specified GSPs of arbitrary complexity.  When creating one's own GSP, it is useful
to keep in mind a handful of requirements for a GSP to be valid. Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder nodes, and $\nonfounds$ for all non-founder
nodes that are not parents of other non-founders, but are adjacent to a sample node, $\nonfounde$ for all
non-founder nodes that have non-founder node daughters, but no sample nodes, and $\nonfoundse$ for all
non-founder nodes that are parents of non-founder nodes and also adjacent to sample-nodes, the necessary conditions
for a valid GSP, in addition to having no inbreeding loops, are:
\begin{equation}
\begin{aligned}
\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders| &  \\
\sum_{i=1}^{e_x} G^-_{x,i} &= 2 & \forall x \in \founders \\
G^+_{x,1} &= G^+_{x,2} & \forall x \in \nonfound \\
\sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\
 S_x/2 + \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfoundse.  \\
\end{aligned}
\end{equation}
Before simulating from a GSP, \gscramble{} checks to ensure the GSP
satisfies these validity conditions.  The simulation of genomic segments
from founders to samples in a GSP is performed by the \gscramble{} function
{\footnotesize\tt segregate()}, which requires as input a list of GSPs through which
to segregate genomic segments, and, for each such GSP, a table that maps the
population labels on each GSP (in the inverted triangles atop the founders) to the
different sampled groups of individuals.

@@

\gscramble{} uses a GSP to segregate segments of genetic material from the founders
to the samples.  At the end of segment simulation, the genome of each sample
is recorded as a mosaic of segments that derived from different founders.  In order
to create a data set of markers in each sample, it is necessary only to identify the alleles
that occurred on each segment within the founders and propagate those alleles to the
corresponding segments within the samples.  At the stage of assigning alleles to
founders, \gscramble{} provides a variety of ways to permute alleles amongst
individuals within samples from which the founders are drawn.  By using permutation,
variability in the data is created, but without sampling with replacement, and hence
not inducing RIPI.

To execute these permutations, it is necessary to have a matrix of genotypes for
the individuals, and also, each population must belong to exactly one
population.  Permutation of alleles then proceeds within populations, hence the frequencies
of alleles from each population remain the same as those observed
in the samples themselves. Following permutation, there remain individuals 1,2,\ldots,
in each population, but the alleles that they carry will be different after permutation.  These
individuals are assigned in order to be the founders from each population in the GSP. Subsequently,
after genetic material is segregated to the samples, the alleles found on each genomic segment in those permuted
founders are then propagated to the samples.  It is worth noting that this approach allows for simulation
with large genomic data sets, with hundreds of thousands to millions of markers, since segregation
through the pedigree is not done via by segments, and not on a locus-by-locus basis. The only operations
involving all the alleles in all the individuals are the permutation steps and indexing into the segments,
which may be done efficiently with matrices in R.

Permutation of alleles is handled within the \gscramble{} function {\footnotesize\tt segments2markers()}
which takes as input the segments segregated down a GSP, a matrix of sampled genotypes, and a specification
of the group that each sampled individuals belongs to.
%%%%%%%
\newcommand{\permcap}{\footnotesize An illustration of how \gscramble{} permutes data under
different options.  {\bf (a)}~A picture
representing an original data set of 12 diploids sampled from three populations and typed at 15 markers. 
Different colors represent different sampled individuals. Individuals 1--5 are shades of blue and green, indicating that they 
are from a single population.  Individuals 6--9 are from another population, (shades of red and orange), while 10--12
are from a different population and are colored shades of violet. Thick vertical black lines separate the different 
populations.  Markers are shown in rows. There are 6 markers on the first chromosome, 5 on the second, and 4 on the 
third.  Each chromosome is delimited by a thick, horizontal black line. The "a" and the "b" in the colored cells denote the first and second gene copies at each marker.  {\bf (b)}~Examples of how the {\tt\footnotesize preserve\_individuals} and {\tt\footnotesize preserve\_haplotypes} options of the {\tt\footnotesize segments2markers()}
function affect the permutation of 
the genetic data of the original data set.  Panel headers show the option values while the colors and "a" and "b" values
in the colored cells show the original position of each gene copy in the original sample.  Note that permutation is always
done within the populations.  When {\tt\footnotesize preserve\_individuals = TRUE},  then just the 
individuals in their entirety are permuted.  When {\tt\footnotesize preserve\_individuals = "BY\_CHROM"}, each pair
of a single chromosome within an individual remains together when permuted. When {\tt\footnotesize preserve\_individuals = FALSE}, gene copies are freely permuted between individuals.  The option {\tt\footnotesize preserve\_haplotypes = TRUE} causes each haplotype on a chromosome to be preserved
while permuting. This option should only be used when the original data set is phased (the phasing is
reflected in the input order of gene copies at each locus).
}
\begin{figure*}
\includegraphics[width=\textwidth]{figures/permutations-plot.png}
\caption[\permcap]{\permcap}
\end{figure*}
%%%%%%%

@@

\gscramble{} REFERENCE GROUPS
\gscramble{} PEDIGREES
ADMIXTURE \gscramble{} VS EMPIRICAL
FST OF EMPIRICAL AND \gscramble{} CORE POPULATIONS
